---
title: "p8105_hw1_bk2959"
author: "Stella Koo"
date: "2024-09-13"
output: github_document
---

## Problem 1
```{r, message = FALSE}
data("penguins", package = "palmerpenguins")

library(tidyverse)
```

The `penguins` dataset contains `r nrow(penguins)` rows and `r ncol(penguins)` columns. 
It provides detailed information about three different `species` of penguins (Adelie, Chinstrap, and Gentoo) and their respective `island` (Biscoe, Dream, and Torgersen). 
The dataset includes important biometric measurements for each penguin, such as:

* Bill length: `bill_length_mm`
* Bill depth: `bill_depth_mm`
* Flipper length: `flipper_length_mm`
* Body mass: `body_mass_g`

The mean flipper length is `r mean(pull(penguins, "flipper_length_mm"), na.rm = TRUE)`.

```{r, warning = FALSE}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()

ggsave("penguins_scatter_plot.pdf")
```

## Problem 2

```{r}
set.seed(1)

problem_df = tibble(
  norm_samp = rnorm(10),
  norm_samp_pos = norm_samp > 0,
  vec_char = c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J"),
  vec_factor = factor(c("Level 1", "Level 1", "Level 2", "Level 3", "Level 2", "Level 1", "Level 2", "Level 3", "Level 3", "Level 1"))
)

problem_df
```

```{r, warning = FALSE}
mean_norm_samp = mean(pull(problem_df, norm_samp))
mean_norm_samp_pos = mean(pull(problem_df, norm_samp_pos))
mean_vec_char = mean(pull(problem_df, vec_char))
mean_vec_factor = mean(pull(problem_df, vec_factor))

mean_norm_samp
mean_norm_samp_pos
mean_vec_char
mean_vec_factor
```


When calculating the mean for each variable, only `norm_samp` and `norm_samp_pos` yielded valid results. This is because `norm_samp` is numeric, and `norm_samp_pos` is logical, where the mean reflects the proportion of TRUE values. In contrast, `vec_char` and `vec_factor` are non-numeric types and results in NA. 

```{r, warning = FALSE}
numeric_logical = as.numeric(pull(problem_df, norm_samp_pos))
numeric_vec_char = as.numeric(pull(problem_df, vec_char))
numeric_vec_factor = as.numeric(pull(problem_df, vec_factor))

mean_numeric_logical = mean(numeric_logical)
mean_numeric_vec_char = mean(numeric_vec_char)
mean_numeric_vec_factor = mean(numeric_vec_factor)
```

After applying `as.numeric` to `norm_samp_pos`, `vec_char`, and `vec_factor`, `norm_samp_pos` and `numeric_vec_factor` yielded valid result when calculating its mean while `numeric_vec_char` produced NA. This occurs because character values such as in `vec_char` cannot be directly converted into numeric form. 

In contrast, `norm_samp_pos` is logical and `as.numeric` converts TRUE values to 1 and FALSE to 0. Similarly, the factor levels in `vec_factor` are converted into their underlying integer codes: "Level 1" becomes 1, "Level 2" becomes 2, and "Level 3" becomes 3.
